### 2018-12-22 tim-lucas-wsc.tex
First pass edits to methods.

The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
To allow for differing relationships between environment and malaria prevalence we fitted separate models for Colombia and Indonesia. 
For Colombia we used all points from South America while for Madagascar we used only Madagascan data.

@@

For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{ranger}, projection pursuit regression \cite{PPR}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
The parameters tried can be seen in Table S1.
Predictions from these models were then made across Colombia and Madagascar.

This model is defined by a likelihood at the level of the polygon (Figure 1A) with covariates and a spatial random field at the pixel-level. Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.

@@

where $\mathrm{p2i}$ is from a model that was published previously. \cite{cameron2015defining} which defines a function
$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$

The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.

@@

The form of the link function means we get predictions of prevalence and incidence simultaneously whether we used both data types or just one.
The linear predictor is composed of an intercept, covariates a spatial, Gaussian random field two iid random effects.

$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v) + w_i(\sigma_w)$$

The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
The first iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$,  was grouped by polygon, with all pixels and point surveys within polygon $j$ being grouped together.
This random effect modelled both missing covariates and extra-Poisson sampling error. 
The second iid random effect, $w_i \sim \operatorname{Norm}(0, \sigma_w)$, was applied to each point survey.
This effect modelled extra-binomial sampling noise.
As such, this random effect is not included in the predicted uncertainty in the incidence or prevalence layers.
However, it is included in predictions of out-of-sample point surveys.

Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$, $\sigma_v$ and $\sigma_w$. We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < \zeta) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. We used different $\zeta$ values for each country: Colombia $\zeta = 1$, Indonesia $\zeta = 3$. We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.

We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}. We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
We assigned $\sigma_w$ a penalised complexity prior such that $P(\sigma_w > 0.3) = 0.0000001$. 
This was chosen by finding the maximum difference in prevalence between survey points (with a sample size greater than 500) within the same raster pixel.
The differences between points within the same pixel can only be accounted for by the binomial error and this iid effect.
Given that the error on a prevalence estimate with sample size greater than 500 is quite small, the iid effect needs to be able to explain this difference.

Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.04)$. 
Given the standardised covariates, a regression coefficient from the 95\% IQR of this distribution, and an intercept of 3, each covariate would be able to predict prevalences between 0.004 and 0.27. 
This prior encodes our belief that the full range of malaria transmission can not be explained by a single covariate and our desire to regularised our model.

The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}

To compare the three models we used two cross validation schemes. 
In the first, polygon incidence data was randomly split into 10 cross-validation folds.
In the second, polygon incidence data was split spatially (5 folds in Colombia, 7 folds in Indonesia).
This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
Our primary performance metric was mean absolute error.

@@

\begin{figure}
\centering
\includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
\caption{{\bf Observed-predicted plots for random cross-validation experiments.}
A) Indonesia, B) Senegal, C) Madagascar. Square root prevalence.
Results from all three models are included with the points only model being shown in orange, the downscaling model being shown in blue and the combined model being shown in green.
}
\label{randompredobspointfacet}
\end{figure}

>>>>

The prevalence survey data were extracted from the Malaria Atlas Project prevalence survey database using only data from, 1990 onwards \cite{bhatt2015effect}. 
As the prevalence surveys cover different age ranges they were standardised to an age range of 2-10 using the model from \cite{smith2007standardizing}. 
For Colombia we used all points from South America while for Madagascar we used only Madagascan data.

@@

For each region we fitted five models via caret \cite{caret}: elastic net \cite{enet}, Random Forest \cite{ranger}, projection pursuit regression \cite{PPR}, neural networks \cite{nnet} and boosted regression trees \cite{gbm}.
Our response variable was prevalence and we weighted the data by sample size (i.e. the number of people tested for malaria in each survey.
For each model we ran 5 fold cross validation to select hyperparameters using random search for Random Forest and boosted regression trees and grid search for the other models. 
%The parameters tried can be seen in Table S1.
Predictions from these models were then made across Colombia and Madagascar.
These predictions were finally inverse logit transformed so that they are on the correct scale.

The top level model is a disaggregation regression model \cite{sturrock2014fine, wilson2017pointless, law2018variational, taylor2017continuous, li2012log}.
This model is defined by a likelihood at the level of the polygon with covariates and a spatial random field at the pixel-level. 
Values at the polygon-level are given the subscript $a$ while pixel level values are indexed with $b$.

@@

where $\mathrm{p2i}$ is from a model that was published previously \cite{cameron2015defining} which defines a function
$${p2i}: f\left(P\right) = 2.616P - 3.596P^2 + 1.594P^3.$$

The fact that the model passes through prevalence space allows predictions of prevalence to be made at the same time as predictions of incidence rate.
Furthermore, it means that the predictions from the machine learning models are on an appropriate scale.
The linear predictor of the model, $\eta_b$, is related to prevalence by a typical logit link function.

@@

The linear predictor is composed of an intercept, covariates, a spatial, Gaussian random and an iid random effects.

$$\eta_b = \beta_0 + \beta X  + u(s, \rho, \sigma_u) + v_j(\sigma_v)$$

The Gaussian spatial effect $u(s, \rho, \sigma_u)$ has a Mat\'ern covariance function and two hyper parameters: $\rho$, the nominal range (beyond which correlation is $< 0.1$) and $\sigma_u$, the marginal standard deviation.
The iid random effect, $v_j \sim \operatorname{Norm}(0, \sigma_v)$, modelled both missing covariates and extra-Poisson sampling error

Finally, we complete the model by setting priors on the parameters $\beta_0, \beta, \rho$ and $\sigma_u$ and $\sigma_v$. We assigned $\rho$ and $\sigma_u$ a joint penalised complexity prior \cite{fuglstad2018constructing} such that $P(\rho < 1) = 0.00001$ and $P(\sigma_u > 1) = 0.00001$. 
We believe that a large proportion of the variance of malaria prevalence and incidence cannot be explained by a linear combination of the covariates selected \cite{bhatt2017improved}, so we set this prior such that the random field could explain most of the range of the data.

We assigned $\sigma_v$ a penalised complexity prior \cite{simpson2017penalising} such that $P(\sigma_v > 0.05) = 0.0000001$. This was based on a comparison of the variance of Poisson random variables, with rates given by the number of cases observed, and an independently derived upper and lower bound for the case counts using the approach defined in \cite{cibulskis2011worldwide}. We found that an iid effect with a standard deviation of 0.05 would be able to account for the discrepancy between the assumed Poisson error and the independently derived error.
Finally, we set regularising priors on the regression coefficients $\beta_i \sim \operatorname{ Norm}(0, 0.4)$. 
The models were implemented and fitted using Template Model Builder \cite{TMB} in R \cite{R}

We compared the performance of the models with three sets of covariates.
Firstly, we used the environmental and anthropogenic covariates, centered and standardised.
Secondly, we used just the predictions from the machine learning models.
Finally we combined these two sets of covariates.

To compare the three models we used two cross validation schemes. 
In the first, polygon incidence data was randomly split into six cross-validation folds.
In the second, polygon incidence data was split spatially into three folds (via k-means clustering on the polygon centroids).
This cross-validation experiment is testing the models’ ability to make predictions far from data where the spatial random field is not informative.
Our primary performance metric was correlation between observed and predicted data.

@@

\begin{figure}
\centering
\includegraphics[trim={0 40mm 0 100mm}, width = 0.6\textwidth]{figs/col_obs_pred_map_ml.png} %\caption{Indonesia spatial crossvalidation} 
\caption{
  Left: Observed polygon data for Colombia. Grey indicates incidence of zero. Right: Out-of-sample predictions for the random cross-validation, machine learning model only model. For each cross-validation fold, predictions are made for the held out data. These predictions are all plotted which together covers the entire study area.
}
\label{randompredobspointfacet}
\end{figure}
